---
title: "Darksky"
output: rmarkdown::html_document
runtime: shiny
---

### Shiny Weather

Dark Sky is an iOS and a website that provides "hyperlocal" weather forecasts. They make their data available to third parties via a web API which we will be using to create a simple shiny app. 

In order to access this API you need an account - if you go to https://darksky.net/dev/ you can sign up for an API account. Once you have registered you will have access to a usage console that includes a unique secret key (the long alphanumeric string at the bottom of the page) you will use to access the API. You can make up to 1000 API requests per day without incurring any cost, so there is no need to enter any billing information.

Documentation for the Dark Sky API can be found [here](https://darksky.net/dev/docs) and includes all information about how to create a properly formated API request and the details of the JSON format of the returned data.


#### Task 1 

In task 1, I wrote a function, `get_darksky`, which allows users to input specific information, `key`, `latitude`, `longtide`, `time` (optional), and the function will download the information and create a data frame.
In the function `get_darksky`, I first write another function `d_darksky` which will generate the url according to the inputs.

Second, I separate the function into 2 parts. The first part will be run if users do not input `date`; therwise, the function will run the second part.

If `date` is NULL, the function will download the `darksky.json` file and create data frame, `df.darksky`.

If `date` is not NULL, the function will download 6 `darksky.joson` files, which contains 2 previous 12 hours data, 1 present data, 3 future 12 hours data. The function tries to get 3 future data is because no matter what date and time we input, darksky will get only 24 hours historical data of the specific date. For example, if we input **2016-12-04 15:00:00**, darksky will report **2016-12-04 00:00:00** to **2016-12-04 23:59:59**. To get full future data, the function therefore download 3 future date from darksky. I then combine each data frame with necessary columns, since with different date, darksky will report different data which will make it difficult to combine raw data frame. In the end, I keep only previous 48 hours, present and next 48 hours in the data frame, `df.darksky`.

Finally, the function will return the data frame `df.darksky` with human time format.

```{r}
library(dplyr)
library(stringr)
library(jsonlite)
library(magrittr)

# `get_darksky` function generates weather data frame
get_darksky = function(key, lat, long, date = NULL) {
  # `d_darksky` function gets the url
  d_darksky = function(key, lat, long, date = NULL) {
    if (is.null(date)) {
      paste0("https://api.darksky.net/forecast/", key, "/", lat, ",", long)
    } else{
      paste0("https://api.darksky.net/forecast/", key, "/", lat, ",", long, ",", date)
    }
  }
  # If input `date` is NULL -> directly download the data from darksky
  if(is.null(date)) {
    # Download the JSON file
    darksky_url = d_darksky(key, lat, long, date)
    download.file(darksky_url, destfile = "darksky.json", quiet = TRUE)
    darksky.json = fromJSON("darksky.json")
    # Generate `df.darksky` data frame
    df.darksky <- data.frame(darksky.json$hourly$data) %>%
      setNames(names(darksky.json$hourly$data)) 
  # If input `date` is not NULL -> obtain previous 48 hours and nest 48 hours data
  # 18 (previous) + 1 (present, input) + 48 (future) = 97 observations
  } else {
    for (i in c(0, -2, -1, 1, 2, 3)) {
      # Download the JOSN file for specific date
      n.date = date + 86400 * i
      darksky_url = d_darksky(key, lat, long, n.date)
      download.file(darksky_url, destfile = "darksky.json", quiet = TRUE)
      darksky.json = fromJSON("darksky.json")
      # i = 0 -> present, generate data frame
      if (i == 0) {
        tmp.darksky <- data.frame(darksky.json$hourly$data) %>%
          setNames(names(darksky.json$hourly$data)) %>%
          .[which(names(darksky.json$hourly$data) %in%
                    c("time", "temperature", "precipProbability", "humidity", "cloudCover", "pressure"))]
      # i != 0 -> previous and future data
      # Combine add-up data with present data frame
      } else {
        tmp.darksky <- rbind(tmp.darksky, data.frame(darksky.json$hourly$data) %>%
                               setNames(names(darksky.json$hourly$data)) %>%
                               .[which(names(darksky.json$hourly$data) %in%
                                         c("time", "temperature", "precipProbability", "humidity", "cloudCover", "pressure"))])
      }
    }
    # Generate `df.darksky` data frame and keep previous and future 48 hours data
    df.darksky <- tmp.darksky[which(tmp.darksky$time >= date - 86400*2 & tmp.darksky$time <= date + 86400* 2), ]
  }
  # Convert the time format, from epoch to human time
  df.darksky$time <- as.POSIXct(df.darksky$time, origin = "1970-01-01")
  return(df.darksky)
}

```


<br/>



#### Task 2

In task 2, the final output would be a tidy data frame `us_page`.

I first read html from wikipedia and get the city table. The table is stored as data frame. Then, I start cleaning the data frame.

First, I remove the footnote number, **[1]**, and unknow symbols, ♠, by using `str_replace` from both column names and the contents.

Second, I convert `estimate` and `census` columns into numeric format by using `str_replace_all` to remove the comma and `as.numeric`. I'll use census column when keeping cities with ore than 500,000 residents in 2010.

Third, I deal with the unit used in `area` and `density` columns. I use `str_extract` and keep only **sq mi**. Then, I rename the column names with (sq ,mi).

Forth, the most complex procedure, I deal with the `Location` column and try to extract latitudes and logitudes from it. When I obtain `location` column, it contains 3 different types of location. I use `str_split_fixed` and keep the one without **N** and **W**. Then, to further clean the data, I keep only numeric part of the latitudes and longitudes and remove the unicodes hidden in the contents and generate `Location` data frame.

After cleaning the data, I combine original data frame, `us_pages`, with latitudes and longitudes data frame, `Location`, and remove the original `location` column.

```{r}
library(rvest)
library(stringr)
library(tibble)
library(dplyr)

# Read html from wikipdeia
page <- read_html("https://en.wikipedia.org/wiki/List_of_United_States_cities_by_population")

# Obtain the city table
us_pages = page %>% 
  html_nodes(xpath = '//*[@id="mw-content-text"]/table[4]') %>% 
  html_table() %>%
  .[[1]]

# Data cleaning -------------------------------------
# Remove the footnote number from the column names
colnames(us_pages) <- str_replace(colnames(us_pages), "\\[[0-9]*\\]", "")
# Remove the footnote number from the table contents
# Remove the unknown symbols from the table contents
col_name = names(us_pages)
for (i in seq_along(col_name)) {
  us_pages[[col_name[i]]] = str_replace(us_pages[[col_name[i]]], "\\[[0-9]*\\]", "")
  us_pages[[col_name[i]]] = str_replace(us_pages[[col_name[i]]], "[0-9]*♠", "")
}
# Convert the content to numeric
conv_name <- c("2015 estimate", "2010 Census")
for (i in seq_along(conv_name)) {
  us_pages[[conv_name[i]]] = str_replace_all(us_pages[[conv_name[i]]], ",", "") %>%
    as.numeric()
}
# Keep only sq mi in area and density columns
us_pages$`2014 land area` = str_extract(us_pages$`2014 land area`, "\\A[0-9]*\\.*[0-9]*") %>%
  str_trim() %>%
  as.numeric()
colnames(us_pages)[which(names(us_pages) == "2014 land area")] <- "2014 land area (sq mi)"
us_pages$`2010 population density` = str_extract(us_pages$`2010 population density`, "\\A[0-9]*\\,*[0-9]*") %>%
  str_trim() %>%
  str_replace_all(",", "") %>%
  as.numeric()
colnames(us_pages)[which(names(us_pages) == "2010 population density")] <- "2010 population density (sq mi)"
# Get the latitudes and longtides from Location column
Location = str_split_fixed(us_pages$Location, "/", 3)[ , 3] %>%
  str_trim() %>%
  str_replace("\\(.*\\)", "") %>%
  str_trim() %>%
  str_extract("\\-*[0-9]*\\.[0-9]*\\;\\s\\-*[0-9]*\\.[0-9]*") %>%
  str_split_fixed(";", 2) %>%
  str_trim() %>%
  as.double() %>%
  matrix(ncol = 2) %>%
  data.frame(stringsAsFactors = FALSE) %>%
  setNames(c("latitudes", "logitudes"))

# Update `us_pages` for tidy latitudes and longtides
us_pages <- cbind(us_pages, Location) %>%
  .[-which(names(us_pages) %in% "Location")] %>%
  .[which(us_pages$`2010 Census` > 500000), ]

```


<br/>
 
#### Task 3 

For the input part, I allow users to input `location`, `date`, `additional information`.

First, I use `selectInput` which allows users to choose one city which is contained the `us_pages` data frame.

Second, I use `checkboxInput` which allows users to decided whether or not to specify a date, it will return as `c_date`.

If the `c_date` is FALSE, the Shiny App will report the prediction of weather. Otherwise, the Shiny App will allow users to input date, which will limit users to choose a date before the present date, and hour.

Third, I use `selectInput` which allows users to choose whether or not to show additional information and what kinds of information they want to see. The options include `Precipitation chance`, `Barometric pressure`, `Humidity` and `Cloud coverage` and will be return as `cond`.

For the output part, I present a summary table and a graph.

First, I get the latitude, `lat`, and logitude, `long`, according to the city users choose. Then, I use `human_date` to get the date and time. If the `c_date` is FALSE, the `human_date` will return the present time; other wise, the `human_date` will return the time chosen by users.

Then, I generate summary table which presents `city`, `state`, `lat`, `long` from the `us_pages` data frame and `human_date` in `output$summary`.

Second, I convert `human_date` into `epoch_date`. The `epoch_date` will be used in the `get_darksky` function later. Note that if the `c_date` is FALSE, the `epoch_date` will return NULL instead of present time.

Third, I input the `lat`, `long` and `epoch` into `get_darksky` function and return the data frame to `darksky`. 

Then, I'm able to generate the graph by `darksky` data frame in `output$temp`.

The basic plot is **temperature**, if the `cond` is `None`. The `temp` plot will have different title depends on whether `c_date` is TRUE or FLASE. If `c_date` is TRUE, the plot title will be **Previous 48 hours & Next 48 hours forcast** and will show an additional vertical line indicates the time specified by users; otherwise, the title will be **Next 48 hours prediction**.

The other information will be added into the plot if the `cond` is not `None`. I put 4 other information as options. The `cond` plot will have different color and node and will present in second y-axis (right y-axis). Then I use function `ggplot_dual_axis` to combine `temp` and `cond` plot and replace the original `temp`. The `output$temp` will return `temp` graph.

```{r echo=FALSE}
library(shiny)
library(shinyTime)
library(dplyr)
library(stringr)
library(scales)
library(parallel)
library(ggplot2)
source("ggplot_dual_axis.R")

shinyApp(
  ui = fluidPage(
    titlePanel("Weather"),
    sidebarLayout(
      sidebarPanel(
        h4("Pick a location and date"),
        hr(),
        # Pick city from us_pages
        selectInput("city", "Cities over 500,000 residents in the States", us_pages$City),
        hr(),
        # Let users choose a date and time
        checkboxInput("c_date", "Choose a specific date and time", value = FALSE),
        # Choose a date (YYYY-MM-DD)
        conditionalPanel(condition = "input.c_date == true",
          # Set max date to input, will be limited to the present time
          dateInput("date", "Date", value = "2016-12-01", min = NULL, max = Sys.Date(),
                    format = "yyyy-mm-dd", startview = "month", weekstart = 0,
                    language = "en", width = NULL),
        # Choose a time (HH:00:00), since darksky uses hour as its unit
        conditionalPanel(condition = "input.c_date == true"),
          selectInput("time", "Hour (military time)", paste0(0:23, ":00:00"))),
        hr(),
        # Let users choose whether ot not to show additional information
        selectInput("cond", "Other information to show",
                    c("None","Precipitation chance", "Barometric pressure", "Humidity", "Cloud coverage")),
        hr(),
        # Copyright
        h6(a("Powered by Dark Sky", href="https://darksky.net/poweredby/"))
      ),
      mainPanel(
        h4("Result"),
        # Summary table which shows the city, state, latitude, longtitute and time the users have entered
        tableOutput("summary"),
        # Plot temperature
        plotOutput("temp")
      )
    )
  ),
  server = function(input, output, session) {
    # Obtain latitudes and longtides according to the city chosen by users
    lat = reactive({
      us_pages$latitudes[which(input$city == us_pages$City)]
    })
    long = reactive({
      us_pages$logitudes[which(input$city == us_pages$City)]
    })
    # Shows the date and time we use to obtain darksky whether data
    human_date = reactive({
      if (input$c_date) {
        return(paste(input$date, input$time))
      } else {
        return(as.character(as.POSIXct(Sys.time())))
      }
    })
    # Convert the date and time to epoch time
    epoch_date = reactive({
      if (input$c_date) {
        return(as.POSIXct(human_date()) %>%
                 as.numeric())
      }
    })
    
    # Input the latitude and longtide and epoch time to get the data frame of weather
    darksky = reactive({
      if (input$c_date) {
        get_darksky("6c3322313fcaf734d302aa4b3be148c9", lat(), long(), epoch_date())
      } else {
        get_darksky("6c3322313fcaf734d302aa4b3be148c9", lat(), long())
      }
    })
    
    # Output summary table which contains the city, state, latitude, longtitute and time the users have entered
    output$summary = renderTable({
      data.frame(input$city, us_pages$State[which(input$city == us_pages$City)],
                 lat(), long(), human_date()) %>%
        setNames(c("City", "State", "latitudes", "longtides", "Date & Time"))
    })
    
    # Output graph which presents the temperature and optional information
    output$temp = renderPlot({
      # Temperature, the scale will be shown in left y axis
      r.temp <- ggplot(data = darksky(), aes(x = time, y = temperature)) +
          geom_line(color = "red") +
          geom_point(color = "red") +
          ylab("Temperature") +
          theme(plot.title = element_text(face = "bold"),
                axis.text.y = element_text(color = "red"),
                axis.title.y = element_text(color = "red"),
                legend.position = "none")
      # Different title to show according to whether or not users specify the date and time
      if(input$c_date) {
        temp <- r.temp + ggtitle("Previous 48 hours & Next 48 hours forcast") +
          geom_vline(xintercept = epoch_date())
      } else {
        temp <- r.temp + ggtitle("Next 48 hours prediction")
      }
      # Additional information: Precipitation change, barometric pressure, humidity, cloud coverage
      # The additional information will be shown in right y axis
      if (input$cond != "None") {
        if (input$cond == "Precipitation chance") {
          cond <- ggplot(data = darksky(), aes(x = time, y = precipProbability)) +
            geom_line(color = "steelblue4") +
            geom_point(color = "steelblue4", shape = 15) +
            scale_y_continuous(labels = percent_format(), limits = c(0,1)) +
            ylab("Precipitation chance (%)") +
            theme(panel.grid = element_blank(),
                  axis.text.y = element_text(color = "steelblue4"),
                  axis.title.y = element_text(color = "steelblue4"),
                  legend.position = "none")
        }
        if (input$cond == "Barometric pressure") {
          cond <- ggplot(data = darksky(), aes(x = time, y = pressure)) +
            geom_line(color = "green4") +
            geom_point(color = "green4", shape = 17) +
            ylab("Barometric pressure") +
            theme(panel.grid = element_blank(),
                  axis.text.y = element_text(color = "green4"),
                  axis.title.y = element_text(color = "green4"),
                  legend.position = "none")
        }
        if (input$cond == "Humidity") {
          cond <- ggplot(data = darksky(), aes(x = time, y = humidity)) +
            geom_line(color = "maroon") +
            geom_point(color = "maroon", shape = 18, size = 2) +
            scale_y_continuous(labels = percent_format(), limits = c(0,1)) +
            ylab("Humidity (%)") +
            theme(panel.grid = element_blank(),
                  axis.text.y = element_text(color = "maroon"),
                  axis.title.y = element_text(color = "maroon"),
                  legend.position = "none")
        }
        if (input$cond == "Cloud coverage") {
          cond <- ggplot(data = darksky(), aes(x = time, y = cloudCover)) +
            geom_line(color = "orange2") +
            geom_point(color = "orange2", shape = 8) +
            scale_y_continuous(labels = percent_format(), limits = c(0,1)) +
            ylab("Cloud coverage (%)") +
            theme(panel.grid = element_blank(),
                  axis.text.y = element_text(color = "orange2"),
                  axis.title.y = element_text(color = "orange2"),
                  legend.position = "none")
        }
        # Out-source package which twist the y axis of additional information to right y axis
        temp <- ggplot_dual_axis(temp, cond, "y")
      }
      return(temp)
    })
  }
)
```